---
title: "Session 9: Bank Marketing - MULTIPLE REGRESSION ANALYSIS"
author: "Lidia Montero"
date: "May 2019"
output: 
  html_document: 
    toc: true
    toc_depth: 3
    number_sections: true
editor_options: 
  chunk_output_type: console
---
# Bank client data

## Description 

*Input variables:*

  1. age (numeric)
  2. job : type of job (categorical: 'admin.','blue-collar','entrepreneur','housemaid','management','retired','self-employed','services','student','technician','unemployed','unknown')
  3. marital : marital status (categorical: 'divorced','married','single','unknown'; note: 'divorced' means divorced or widowed)
  4. education (categorical:'basic.4y','basic.6y','basic.9y','high.school','illiterate','professional.course','university.degree','unknown')
  5. default: has credit in default? (categorical: 'no','yes','unknown')
  6. housing: has housing loan? (categorical: 'no','yes','unknown')
  7. loan: has personal loan? (categorical: 'no','yes','unknown')# related with the last contact of the current campaign:
  8. contact: contact communication type (categorical:'cellular','telephone')
  9. month: last contact month of year (categorical: 'jan', 'feb', 'mar',..., 'nov', 'dec')
  10. day_of_week: last contact day of the week (categorical:'mon','tue','wed','thu','fri')
  11. duration: last contact duration, in seconds (numeric). Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no'). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.
  12. campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
  13. pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
  14. previous: number of contacts performed before this campaign and for this client (numeric)
  15. poutcome: outcome of the previous marketing campaign (categorical: 'failure','nonexistent','success')# social and economic context attributes
  16. emp.var.rate: employment variation rate - quarterly indicator (numeric)
  17. cons.price.idx: consumer price index - monthly indicator (numeric)
  18. cons.conf.idx: consumer confidence index - monthly indicator (numeric)
  19. euribor3m: euribor 3 month rate - daily indicator (numeric)
  20. nr.employed: number of employees - quarterly indicator (numeric)
  21. y - has the client subscribed a term deposit? (binary: 'yes','no')

```{r}
# Load Required Packages: to be increased over the course

requiredPackages <- c("mvoutlier","chemometrics","mice","missForest","missMDA","DMwR","pbkrtest","jomo","readxl","haven","sf","rgdal","effects","FactoMineR","car","factoextra","RColorBrewer","ggplot2","dplyr","data.table", "ggmap","ggthemes","knitr","MVA","car")
missingPackages <- requiredPackages[!(requiredPackages %in% as.vector(installed.packages(lib.loc="~/R/win-library/3.5")[,"Package"]))]

#if(length(missingPackages)) install.packages(missingPackages)
#install.packages(requiredPackages,dependencies = TRUE,repos = c("http://rstudio.org/_packages","http://cran.rstudio.com"))
lapply(requiredPackages, require, character.only = TRUE)

```

## Load your sample of 5000 records already cleaned and with all variables properly coded

```{r}
rm(list=ls())
# Load Required Packages: to be increased over the course

#setwd("C:/Users/lmontero/Dropbox/DOCENCIA/FIB-ADEI/PRACTICA/BankMarketing")
setwd("D:/DOCENCIA/FIB-ADEI/PRACTICA/BankMarketing/LABS")
#file_path <- "D:/Dropbox/DOCENCIA/FIB-ADEI/PRACTICA/BankMarketing/" 
load("MyBank5000.RData")
summary(df)
```

# Explanatory numeric variables

## Initial modelling

```{r}
# Guillem - 
vars_con
#vars_con<-names(df)[c(1,12:14,16:20)]
m1<-lm(duration~.,data=df[,c("duration",vars_con)])
summary(m1)

# Inferential criteria o Bayesian info criteria
# Remove non significant variables
library(car)
Anova(m1)  # Mètode testar els efectes nets
m2<-lm(duration~campaign+pdays+previous+emp.var.rate+nr.employed,data=df)
Anova(m2)  # Test de Fisher
# Akaike or BIC
m3<-step(m1)  # AIC
m4<-step(m1,k=log(nrow(df))) # BIC Large data sets

# Alternativa no Guillem - Recomanable quan hi ha moltes variables

# Usa condes()
names(df)
condes(df,11)
#  previous+emp.var.rate+euribor3m+campaign+pdays+nr.employed

m5<-lm(duration~previous+emp.var.rate+euribor3m+campaign+pdays+nr.employed,data=df)
Anova(m5) # euribor3m net effect non significant
m6<-lm(duration~previous+emp.var.rate+campaign+pdays+nr.employed,data=df)
Anova(m6) 

# BIC
m7<-step(m5,k=log(nrow(df)))

# Both results m4 and m7 are equal (in my case)
vif(m7)

```

## Transforming variables

```{r}
par(mfrow=c(2,2))
plot(m7)
par(mfrow=c(1,1))

# Box-Cox Transformation of Y - log(Y)
library(MASS)
boxcox(m7,data=df)  # lambda 0 -> log(Y)

m8<-lm(log(duration)~previous+emp.var.rate+euribor3m+campaign+pdays+nr.employed,data=df)
Anova(m8) # euribor3m, previous net effects non significant
m9<-lm(log(duration)~emp.var.rate+campaign+pdays+nr.employed,data=df)
Anova(m9) 

# BIC
m10<-step(m8,k=log(nrow(df)))

# emp.var.rate and nr.employed are showing large colinearity
vif(m10) # vif >3 dangerous -> colinearity

m11<-lm(log(duration)~emp.var.rate+campaign+pdays,data=df)
summary(m11)

# Polinomic regression

m20<-lm(log(duration)~poly(emp.var.rate,2)+poly(campaign,2)+poly(pdays,2),data=df)
summary(m20)
Anova(m20)
marginalModelPlots(m11)

```


